---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Loading additional libraries:
```{r results="hide"}
library(dplyr)
library(ggplot2)
```

```{r}
setwd("C:\\Users\\hp pc\\RepData_PeerAssessment1")
data<-read.csv("activity/activity.csv")
```

## The following plot shows the total number of steps taken each day.
```{r}
group_data<-group_by(data,date)
summ_data<-summarize(group_data,steps=sum(steps,na.rm=TRUE))
hist(summ_data$steps,col="blue",xlab="Number of steps",main="Histogram of total number of steps")
```

## Mean and median number of steps taken each day
```{r}
mn=mean(summ_data$steps,na.rm=T)
md=median(summ_data$steps,na.rm=T)
```
The mean number of steps taken each day is `r mn` and the median number of steps each day is `r md`

## Time series plot of the average number of steps taken
```{r}
group_data<-group_by(data,interval)
summ_data<-summarize(group_data,steps=mean(steps,na.rm=TRUE))
with(summ_data,plot(interval,steps,type="l",xlab="Five minute intervals",main="Time series plot of the average number of steps taken"))
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
m<-max(summ_data$steps)
minterval<-as.numeric(select(filter(summ_data,steps==m),interval))
```

The 5-minute interval that,on average,contains the maximum numbe of steps is `r minterval`.

## Code to describe and show a strategy for imputing missing data
```{r}
nmiss<-sum(is.na(data$steps))
```

The number of rows containing NA values are `r nmiss` 

### imputing missing values
```{r}
id<-is.na(data$steps)
data[id,1]<-mean(data$steps,na.rm=TRUE)
```
The missing values have been imputed with the mean of the steps.

## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
group_data<-group_by(data,date)
summ_data<-summarize(group_data,steps=sum(steps,na.rm = TRUE))
hist(summ_data$steps,xlab="Number of steps",main="Total number of steps",col="gray")
mtext("The missing values have been imputed with the mean",outer=FALSE)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
data<-transform(data,date=as.Date(date))
weekend<-c("Saturday","Sunday")
data$day_type<-ifelse(weekdays(data$date) %in% weekend,"weekend","weekday")
group_data<-group_by(data,day_type,interval)
summ_data<-summarize(group_data,steps=mean(steps,na.rm=TRUE))
ggplot(summ_data,aes(interval,steps))+geom_line()+facet_grid(.~day_type)
```

